---
title: "Untitled"
format: html
---

```{r}
#to clear objects from R environment
rm(list = ls())
```


```{r}
#to find my working directory 
getwd()
```


```{r}
# Load library
library(dplyr)

# Set your working directory to the folder with the CSVs
setwd("C:/Users/sarao/OneDrive - St. Lawrence University/Summer Research/summer research/C.elegans/data/data_original")

# Get list of all CSV files
csv_files <- list.files(pattern = "\\.csv$", full.names = TRUE)

# Read and combine all CSV files into one big dataframe
combined_data <- lapply(csv_files, read.csv) %>%
  bind_rows()

write.csv(combined_data, "data_original.csv", row.names = FALSE)


```

### Problem
I discovered that some of my .csv files contain misspelled column headers, which caused extra columns to appear and misaligned my data across columns. I now need to locate these files and correct the errors. The misspelled headers include "aagr_area", "X", "Area", and "worm_ID".


```{r}
# Set working directory
setwd("C:/Users/sarao/OneDrive - St. Lawrence University/Summer Research/summer research/C.elegans/data/data_original")

# List CSVs
csv_files <- list.files(pattern = "\\.csv$", full.names = TRUE)
target_cols <- c("aagr_area", "X", "Area", "worm_ID")

# Initialize result lists
column_check <- list()
skipped_files <- c()

for (file in csv_files) {
  tryCatch({
    # Read only first few lines
    df <- read.csv(file, nrows = 5)
    matched_cols <- intersect(names(df), target_cols)
    
    if (length(matched_cols) > 0) {
      column_check[[file]] <- paste(matched_cols, collapse = ", ")
    }
  }, error = function(e) {
    # Record skipped files
    skipped_files <<- c(skipped_files, basename(file))
  })
}

# Print files that matched
if (length(column_check) > 0) {
  cat("✅ Files containing matching columns:\n")
  for (f in names(column_check)) {
    cat(basename(f), "→", column_check[[f]], "\n")
  }
} else {
  cat("❌ No files matched the column names.\n")
}

# Print skipped files
if (length(skipped_files) > 0) {
  cat("\n⚠️ Skipped unreadable/empty files:\n")
  print(skipped_files)
}

```

Files containing matching columns:
C3_N3_D1.csv → aagr_area 
combined_data.csv → aagr_area, X, Area, worm_ID 
M2_N1_D1.csv → X, Area 
M3_N2_D3.csv → worm_ID 

### Solution
Now I will go back into those .csv files, fix the column names, and re upload them to the data folder.

```{r}
# creating the dataset I will use for my analysis

library(dplyr)

# Define paths relative to your current working directory or absolute paths
data_fixed_path <- "C:/Users/sarao/OneDrive - St. Lawrence University/Summer Research/summer research/C.elegans/data/data_fixed"
dataset_final_path <- "C:/Users/sarao/OneDrive - St. Lawrence University/Summer Research/summer research/C.elegans/data/dataset_final"

# Get list of all CSV files in data_fixed
csv_files <- list.files(path = data_fixed_path, pattern = "\\.csv$", full.names = TRUE)

# Read and combine all CSV files
celegan_data <- lapply(csv_files, read.csv) %>%
  bind_rows()

# Write combined data to dataset_final folder
write.csv(celegans_data, file = file.path(dataset_final_path, "dataset_final.csv"), row.names = FALSE)

#did not save with the name I wanted so I manually renamed the data set 
```

